Computing with cut systems +

### 2015-10-14 cutsystems-new.tex
Revisions.

% TODO
% - fix the equivalence of L with LG
% - add stuff to the example

@@

\begin{abstract}
A cut system has objects, $1$-morphisms and $2$-morphisms, similar to a bicategory. In contrast to the composition of $1$-morphisms in a bicategory, the cut operation on $1$-morphims in a cut system produces representations of Clifford algebras in supercategories rather than bare objects. The motivating example of a cut system refines the bicategory of Landau-Ginzburg models, in which the objects are hypersurface singularities and the $1$-morphisms are matrix factorisations.
\end{abstract}

@@

The bicategory of Landau-Ginzburg models $\LG$ has polynomials $W(x)$ called \textsl{potentials} as its objects, and $1$-morphisms $W(x) \lto V(y)$ are matrix factorisations of the difference $V(y) - W(x)$ over $\nC[x,y]$. This bicategory has been studied extensively in connection with topological field theory \cite{??} and knot invariants \cite{??}. The most important and nontrivial structure in this bicategory is the functor
\be\label{eq:intro_functor_comp}
\LG(V,U) \otimes_{\nC} \LG(W, V) \lto \LG(W,U)
\ee
which composes $1$-morphisms for potentials $W(x),V(y),U(z)$. Continuing the approach developed with Dyckerhoff in \cite{??}, the aim of this paper is to explicitly describe this functor. Any increase in our understanding of the operation of composition in $\LG$ has implications for the subjects where this bicategory makes an appearance, but our primary application is to the semantics of linear logic constructed in \cite{??}.

\begin{itemize}
\item We define another bicategory without units $\cat{L}$
\item We prove $\cat{L}$ is equivalent to $\LG$
\item We show how to model $\cat{L}$ in schemes (next paper)
\end{itemize}

First of all, let us be more precise about the kind of explicit description we are looking for. Observe that once we have chosen a homogeneous basis, a matrix factorisation of $W(x) - V(y)$ is just a matrix of polynomials
\[
D \in M_{2r}( \nC[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (W(x) - V(y)) \cdot I_{2r}\,.
\]
If we fix the rank $r$ and bound the degrees of the monomials appearing in $D$ by some integer $k$, it is easy to parametrise such matrix factorisations by the closed points of an affine scheme $\mathscr{M}_{r,k}(W,V)$. Then the set of objects of $\LG(W,V)$ can be described as a disjoint union of inductive limits of these schemes
\[
\operatorname{ob} \LG(W,V) = \bigsqcup_{r \ge 1} \bigcup_{k \ge 0} \mathscr{M}_{r,k}(W,V)\,.
\]
The kind of explicit description we are looking for is a lifting of the function \eqref{eq:intro_functor_comp} on the level of objects to a morphism of schemes
\be\label{eq:composition_as_schememap}
\mathscr{M}_{r',k'}(V,U) \times \mathscr{M}_{r,k}(W,V) \lto \mathscr{M}_{f(r,r'),g(k,k')}(W, U)\,.
\ee
Unfortunately this appears difficult, if not impossible, as stated. In this article we explain how the replacement of $\LG(W,U)$ by an equivalent category $\cat{L}(W,U)$, called the \textsl{Clifford thickening}, allows for such a lifting of \eqref{eq:intro_functor_comp} to a morphism of schemes.

% One consequence is that $\LG$ is equivalent to the bicategory $\cat{L}^\bullet$, which is \emph{computable} in the sense that its operations may be embodied in software. To some extent this software already exists: a component of what we are here calling the cut system $\cat{L}$ (developed earlier in joint work with Dyckerhoff \cite{dm1102.2957}\footnote{For a more detailed discussion of the relation of this work with the present paper, see Remark \ref{remark:relation_to_toby_paper}.}) has already been implemented in the computer algebra package Singular in joint work with Carqueville \cite{khovhompaper}, and used to compute tables of Khovanov-Rozansky knot homologies. The code was also used in the discovery of the generalised orbifolding relations among ADE singularities by Carqueville, Ros Camacho and Runkel \cite{ade}. 
\medskip

@@

@@

\emph{Acknowledgements.} Thanks to Nils Carqueville and Ingo Runkel for helpful comments on the draft. Although the sequel on linear logic is the proper place to fully acknowledge the debt, we cannot neglect to mention the important role that the beautiful ideas in Girard's geometry of interaction program \cite{girard_towards} had in shaping the present work.

@@

Let $R = k[x_1,\ldots,x_n]$ be a polynomial ring and $X$ a finite rank matrix factorisation of $W$ over $R$. Let us fix a homogeneous basis for $X$ so that we may identify $d_X$ with a matrix, and differentiate this matrix entry-wise to obtain the matrix $\lambda_i = \partial_{x_i}(d_X)$.

\begin{lemma} The odd $R$-linear operator $\lambda_i$ on $X$ is independent, up to homotopy, of the choice of basis.
\end{lemma}

@@

\begin{lemma}\label{lemma:naturalityoflambda} Let $\phi: X \lto X'$ be a morphism of matrix factorisations and write $\lambda_i = \partial_{x_i}(d_X)$ and $\lambda'_i = \partial_{x_i}(d_{X'})$. Then there is a homotopy $\phi \lambda_i \simeq \lambda'_i \phi$.
\end{lemma}
\begin{proof}
From $d_{X'} \phi = \phi d_X$ we deduce
\[
\partial_{x_i}( d_{X'} ) \phi + d_{X'} \partial_{x_i}( \phi ) = \partial_{x_i}( \phi ) d_X + \phi \partial_{x_i}( d_X )
\]
and hence
\[
\partial_{x_i}( d_{X'} ) \phi - \phi \partial_{x_i}( d_X ) = \partial_{x_i}(\phi) d_X - d_{X'} \partial_{x_i}(\phi)
\]
as claimed.
\end{proof}

@@

In a cut system the composition rule produces objects which are representations of Clifford algebras. We briefly recall the basic theory of these algebras and their modules; for a full discussion see \cite{friedrich}. For $n \ge 0$ the Clifford algebra $C_n$ is the associative $\mathbb{Z}_2$-graded $k$-algebra generated by odd elements $\gamma_1,\ldots,\gamma_n$ and $\gamma_1^\dagger, \ldots, \gamma_n^\dagger$ subject to the following relations
\be\label{eq:clifford_relations}
[\gamma_i, \gamma_j] = 0, \qquad [\gamma_i^\dagger, \gamma_j^\dagger] = 0, \qquad [\gamma_i, \gamma_j^\dagger] = \delta_{ij}\,.
\ee
for all $1 \le i, j \le n$. In this paper all our commutators are graded, $[a,b] = ab - (-1)^{|a||b|} ba$. We set $C_0 = k$. These Clifford algebras have essentially only one nontrivial representation, sometimes called the \emph{spinor representation} \cite[p.14]{friedrich}. To define this representation we use the $\nZ_2$-graded $k$-module
\[
F_n = k \theta_1 \oplus \cdots \oplus k \theta_n
\]
where the $\theta_i$ are formal variables of odd degree, and set
\be
S_n = \bigwedge F_n = \bigwedge\big( k \theta_1 \oplus \cdots \oplus k \theta_n \big)\,.
\ee

@@

In particular this means that every $\mathbb{Z}_2$-graded $C_n$-module is isomorphic to $S_n \otimes_k V$ for some $\mathbb{Z}_2$-graded $k$-module $V$. Using the isomorphism of $k$-modules
\[
F_m \oplus F_n \cong F_{m+n}
\]
which maps the ordered basis $\theta_1,\ldots,\theta_m$ of $F_m$ to the first $m$ basis elements of $F_{m+n}$ in their given order, and the ordered basis of $F_n$ to the last $n$ basis elements, we define
\[
S_m \otimes_k S_n = \bigwedge F_m \otimes_k \bigwedge F_n \cong \bigwedge F_{m+n} = S_{m+n}\,.
\]
From this we deduce an isomorphism of algebras
\begin{equation}\label{eq:algebra_A_additive}
C_m \otimes_k C_n \cong C_{m+n}\,.
\end{equation}
Given $m, n \ge 0$ we introduce notation for the $C_m$-$C_n$-bimodule
\begin{equation}
S_{m,n} = S_m \otimes_k S_n^* = \Hom_k(S_n,S_m)
\end{equation}
where $S_n^* = \Hom_k(S_n, k)$. There is an isomorphism of $C_l$-$C_n$-bimodules
\begin{equation}\label{eq:isosbimodule}
S_{l,m} \otimes_{A_m} S_{m,n} = S_l \otimes_k S_m^* \otimes_{C_m} S_m \otimes_k S_n^* \cong S_{l,n}\,.
\end{equation}

@@

\section{The Clifford thickening}

@@

More concretely $\cat{T}^\bullet$ is the category with objects all pairs $(X,n)$ of an integer $n \ge 0$ and an object $X \in \cat{T}$ together with a left $C_n$-action. A morphism $\alpha: (X,n) \lto (Y,m)$ is a morphism of $C_m$-modules $\alpha: S_{m,n} \otimes_{C_n} X \lto Y$. Composition of a pair of morphisms
\begin{equation}\label{eq:comp_chain0}
\xymatrix{
(X,n) \ar[r]^-{\alpha} & (Y,m) \ar[r]^-{\beta} & (Z,l)
}
\end{equation}
is given by the morphism of $C_l$-modules
\begin{equation}\label{eq:comp_chain1}
\xymatrix@C+1pc{
S_{l, n} \otimes_{C_n} X \ar[r]^-{\cong} & (S_{l,m} \otimes_{C_m} S_{m,n}) \otimes_{C_n} X \ar[d]^-{\cong}\\
& S_{l,m} \otimes_{A_m} ( S_{m,n} \otimes_{C_n} X ) \ar[r]^-{ 1 \otimes \alpha } & S_{l,m} \otimes_{C_m} Y \ar[r]^-{\beta} & Z\,.
}
\end{equation}
There is a canonical functor $p: \cat{T}^\bullet \lto \mathbb{N}$ defined by $p(X,n) = n$, which defines a cofibered category over $\mathbb{N}$ by the definition of the Grothendieck construction. This definition has the benefit of being derived from a general construction, but there is an alternative description of morphisms which is less cumbersome:

@@

In this section we define a bicategory without units which we call the \textsl{cut system} model of $\LG$. Later in Section \ref{??} we will prove that this bicategory is equivalent to $\LG$, although this will not be apparent at first. Given objects $(x,W)$ and $(y,V)$ of $\LG$ we define
\be\label{eq:defnluv}
\cat{L}(W,V) = \Big( \hmf( k[x,y], V - W )^{\omega} \Big)^{\bullet}\,,
\ee
where $(-)^\omega$ denotes the idempotent completion and $(-)^{\bullet}$ the Clifford thickening. The first aim of this section is to define, for any object $(z,U)$ of $\LG$, a functor
\begin{gather*}
\cat{L}(V,U) \otimes_k \cat{L}(W,V) \lto \cat{L}(W, U)\,\\
(Y,X) \mapsto Y \l X
\end{gather*}
which we call the \textsl{cut functor}. The cut operation is defined on matrix factorisations $X$ of $V - W$ and $Y$ of $U - V$ as follows, assuming $y = (y_1,\ldots,y_m)$ and writing
\be\label{eq:defnjacobian}
J_V = k[y] / ( \partial_{y_1} V, \ldots, \partial_{y_m} V )\,.
\ee

@@

>>>>

@@

\begin{abstract}
In the bicategory of Landau-Ginzburg models $\LG$ the objects are certain polynomials called potentials, and the $1$-morphisms are matrix factorisations. The composition of these $1$-morphisms is defined by a form of convolution, and is not easy to compute. In this paper we define a bicategory $\cat{L}$ which is equivalent to $\LG$ but for which composition of $1$-morphisms is given explicitly. This construction will be applied in a sequel to define an enrichment of $\LG$ over affine schemes.
\end{abstract}

@@

The bicategory of Landau-Ginzburg models $\LG$ has polynomials $W(x)$ called \textsl{potentials} as its objects (typically, the defining equation of an isolated singularity) and the $1$-morphisms $W(x) \lto V(y)$ are matrix factorisations of the difference $V(y) - W(x)$ over $\nC[x,y]$. This bicategory plays an important role in topological field theory \cite{McNameethesis,lgdual,genorb} and the study of homological knot invariants \cite{kr0401268}. In \cite{lgdual} it was shown that $\LG$ is a pivotal bicategory with adjoints, and closed formulas were given for the units and counits of adjunction. In this paper and its sequel we study an enrichment of $\LG$ which we refer to as a \emph{cut system}.

A finite rank matrix factorisation of $V(y) - W(x)$ is a matrix of polynomials
\[
D \in M_{2r}( \nC[x,y] ) \qquad \text{ satisfying } \qquad D^2 = (V(y) - W(x)) \cdot I_{2r}
\]
where we require that, dividing $D$ into $r \times r$ blocks, the upper left and bottom right blocks are zero matrices. The set of finite rank matrix factorisations is denoted $\LG(W,V)_{fin}$. For technical reasons the category of $1$-morphisms $\LG(W,V)$ contains, in addition to the finite rank matrix factorisations, those infinite rank matrix factorisations which are homotopy equivalent to finite rank ones; this means in particular that the inclusion
\[
\LG(W,V)_{fin} \lto \LG(W,V)
\]
is an equivalence of categories.

If we fix the size $r$ of the blocks and bound the degrees of the monomials appearing in $D$ by some integer $k$, we can parametrise the possible matrix factorisations by the closed points of an affine scheme $\mathscr{M}_{r,k}(W,V)$ whose coordinates are the coefficients of the various monomials in each row and column of $D$. Hence as sets we have
\begin{equation}\label{eq:param_frmf}
\LG(W,V)_{fin} = \bigsqcup_{r \ge 1} \bigcup_{k \ge 0} \mathscr{M}_{r,k}(W,V)_{\nC}\,.
\end{equation}
For various applications, two of which we sketch below, we would like to know that the basic operations of the bicategory $\LG$ respect this additional scheme structure; that is, that these operations are continuous in the coordinates which parametrise the $1$-morphisms. The most important and nontrivial such structure is the composition functor
\begin{gather*}
\LG(V,U)_{fin} \times \LG(W, V)_{fin} \lto \LG(W,U)\,,\\
(Y,X) \mapsto Y \otimes X\,.
\end{gather*}
We would like therefore to produce a morphism of schemes (omitting indices)
\[
c: \mathscr{M}(V,U) \times \mathscr{M}(W,V) \lto \mathscr{M}(W,U)
\]
with the property that $c([Y], [X]) = [Y \otimes X]$, where $[-]$ denotes the point of the parameter space corresponding to a given matrix factorisation.

The problem is that this doesn't make any sense, because $Y \otimes X$ contains monomials $x^i y^j z^k$ with $y$'s in them (we say $Y \otimes X$ is \emph{infinite rank} over $k[x,z]$) and we have no way to store the coefficients of such monomials in the space $\mathscr{M}(W,U)$. However one can prove that $Y \otimes X$ is a summand in the homotopy category of a finite rank matrix factorisation (involving only $x$'s and $z$'s) and even describe this object and the corresponding idempotent explicitly \cite{dm1102.2957}. This is a finite amount of data which can be used to encode $Y \otimes X$, so we just need to replace $\mathscr{M}(W,U)$ by some scheme which knows how to store such data.

\vspace{0.5cm}

After some refinements\footnote{See Remark \ref{remark:relation_to_toby_paper} for a discussion of why we end up encoding the finite model by Clifford actions rather than the idempotents of \cite{dm1102.2957}.} this amounts to replacing $\LG$ by an equivalent bicategory $\cat{L}$ with the same objects and (almost) the same $1$-morphisms, but whose composition operation is easily encoded as a morphism of schemes. We say \emph{almost} the same $1$-morphisms because in $\cat{L}$ the $1$-morphisms are still finite rank matrix factorisations, but now they may additionally equipped with the action of a Clifford algebra. It remains straightforward to parametrise these $1$-morphisms by the closed points of an affine scheme, but in contrast to $\LG$ the composition functor in $\cat{L}$ (which we refer to as the \emph{cut operation})
\begin{gather}
\cat{L}(V,U) \times \cat{L}(W,V) \lto \cat{L}(W,U)\,,\\
(Y,X) \mapsto Y \l X
\end{gather}
does indeed lift to a morphism of the parametrising schemes - fulfilling the original goal. The schemes parametrising $1$-morphisms in $\cat{L}$ together with the morphisms of schemes lifting composition, and vector bundles encoding the $2$-morphisms, is what we refer to as a \emph{cut system} in the sequel \cite{cutsystems2}. However the present paper is concerned only with defining the bicategory $\cat{L}$ and proving that it is equivalent to $\LG$.

@@

\subsection{Applications}

\subsubsection{Topological field theory}

The bicategory $\LG$ arises in topological field theory (TFT) as the data necessary to define a two-dimensional topological field theory with defects \cite{k1004.2307,dkr1107.0495}. It is expected that such a TFT should be associated to any suitably nice bicategory $\cat{B}$. In the formalisation of \cite[\S 3.1]{dkr1107.0495} the associated TFT is a symmetric monoidal functor
\[
\mathcal{Z}: \operatorname{Bord}_{2,1} \lto \operatorname{Vect}_\nC
\]
where the objects of $\operatorname{Bord}_{2,1}$ are disjoint unions of (collared) circles labelled with objects and $1$-morphisms of $\cat{B}$, and the morphisms are diffeomorphism equivalence classes of two-dimensional manifolds with boundary and a labelled network of ``defects''.

\begin{figure}[bt]
$$
  \raisebox{-75pt}{\begin{picture}(300,130)
  \put(0,0){\scalebox{0.6}{\includegraphics{pic02.eps}}}
  \put(0,0){
     \setlength{\unitlength}{.55pt}\put(-22,-8){
%     \put(136,224)   {$ \phi_\text{in} $}
%     \put(140, 22)   {$ \phi_\text{in} $}
%     \put(433,142)   {$ \phi_\text{out} $}
	 \put(295,85)   {$ Y $}
	 \put(240,190)   {$ X $}
     \put(510,180)   {$ O_{3} $}
     \put( 10,230)   {$ O_{1} $}
     \put( 10,110)   {$ O_{2} $}
     }\setlength{\unitlength}{1pt}}
  \end{picture}}
$$
\caption{A world-sheet $M$ with defect lines from $\,(O_1,O_2)\,$ to
$\,(O_{3})$.\ The two-dimensional regions of the world-sheet are all labelled with the same potential $W(x)$, while $X,Y$ are matrix factorisations of $W(x) - W(x')$. The labels on the other defect lines are omitted.}
\label{fig:wsh-defect}
\end{figure}

Consider the example shown in Figure \ref{fig:wsh-defect}, where $\cat{B} = \LG$, the two-dimensional regions are all labelled with a single potential $W(x)$ and $X,Y: W \lto W$ are $1$-morphisms. The value of $\mathcal{Z}$ on $M$ is a $\nC$-linear map
\be\label{eq:value_of_bordism}
\mathcal{Z}(M): \mathcal{Z}(O_1) \otimes \mathcal{Z}(O_2) \lto \mathcal{Z}(O_3)\,.
\ee
We would like to discuss an additional property of the physicist's Landau-Ginzburg model which is not encoded into the data of the functor $\mathcal{Z}$. Namely, it is a consequence of the way the defect conditions $X,Y$ enter the partition function \cite[(3.16)]{laz} that the vector spaces $\mathcal{Z}(O_i)$ and linear map $\mathcal{Z}(M)$ should depend continuously on the coefficients defining $Y,X$. To represent this mathematically we might try to present $\mathcal{Z}(O_i)$ as a vector bundle over a scheme $\mathscr{M}_Y \times \mathscr{M}_X$ parametrising possible labels and $\mathcal{Z}(M)$ as a morphism of these vector bundles. But there are some obstacles.
% \cite[(2.9)]{dkr1107.0495}

Consider the incoming circle $O_2$ with its labels $Y,X$. The construction gives
\be\label{eq:zo2orig}
\mathcal{Z}(O_2) := H^0 \Hom_{\nC[x,x']}( Y, X )
\ee
which is the vector space of $2$-morphisms $Y \lto X$ in $\LG(W,W)$. Now suppose that the defect conditions are made to depend on parameters 
\[
Y = Y(t_1,\ldots,t_r)\,, \qquad X = X(u_1,\ldots,u_s)
\]
which satisfy their own equations and cut out a parametrising scheme $\mathscr{M}_Y \times \mathscr{M}_X \subseteq \mathbb{A}^{r+s}$. The question is: what kind of global object $\mathcal{Z}(O_2)$ on $\mathscr{M}_Y \times \mathscr{M}_X$ is formed by the finite-dimensional vector spaces
\[
\mathcal{Z}(O_2)_{P,Q} := H^0 \Hom_{\nC[x,x']}( Y(P), X(Q) )
\]
as $P$ varies over $\mathscr{M}_Y$ and $Q$ over $\mathscr{M}_X$? The problem is that \emph{a priori} we have
\be\label{eq:cohom_sucks}
\big\{ H^* \Hom_{\nC[x,x',t,u]}( Y(t), X(u) ) \big\}_{P,Q} \neq H^* \Hom_{\nC[x,x']}( Y(P), X(Q) )\,.
\ee
%This is a particular case of a general phenomenon: any complex of $\nC$-vector spaces $C$ is isomorphic in the derived category $\mathbb{D}(\nC)$ to $\bigoplus_n H^n(C)[n]$, but over a general ring, for instance the coordinate ring of $\mathscr{M}_Y \times \mathscr{M}_X$, this is not the case. 
The solution is to work instead with the full complex, which suggests the definition\footnote{Note that it is not enough to take the derived category of coherent sheaves, because the Hom complex is made up of non-finitely-generated modules over $\mathscr{M}_Y \times \mathscr{M}_X$.}
\be\label{eq:wuhe20}
\mathcal{Z}(O_2) := \Hom_{\nC[x,x',t,u]}( Y(t), X(u) ) \in \mathbb{D}( \operatorname{qcoh} \mathscr{M}_Y \times \mathscr{M}_X )\,.
\ee
An alternative (and we believe simpler) approach is to take $\cat{L}$ rather than $\LG$ as the input to the definition of the TFT. Before doing so, let us observe that in the situation where $X,Y$ do not depend on parameters, we may equivalently\footnote{This definition is not quite equivalent, because the displayed composite produces a complex of vector spaces quasi-isomorphic to $H^* \Hom(Y,X)$ rather than $H^0 \Hom(Y,X)$, but this is not important here.} define $\mathcal{Z}(O_2)$ to be the following composite in $\LG$:
\be\label{eq:composite_intro_o2}
\xymatrix@C+2pc{
0 \ar[r]^-X & W(x) - W(x') \ar[r]^-{Y^{\vee}} & 0
}\,.
\ee
Now let $\mathcal{Z}'$ be the TFT defined using the bicategory $\cat{L}$ as input. Then $\mathcal{Z}'(O_2)$ may likewise be computed by the composite \eqref{eq:composite_intro_o2} in $\cat{L}$, which is a $\nZ_2$-graded complex of vector spaces $Y^{\vee} \l X$ equipped with the action of a Clifford algebra. If $Y(t), X(u)$ are allowed to depend on parameters, the cut becomes a vector bundle
\be\label{eq:orynx}
\mathcal{Z}'(O_2) := Y^{\vee}(t) \l X(u)
\ee
over $\mathscr{M}_Y \times \mathscr{M}_X$ equipped with a differential and compatible Clifford action, such that
\be
\{ Y^{\vee}(t) \l X(u) \}_{P,Q} \cong Y^{\vee}(P) \l X(Q)
\ee
for every point $(P,Q)$.

Since $\cat{L}$ and $\LG$ are equivalent, the two ways \eqref{eq:wuhe20},\eqref{eq:orynx} of reading the composite \eqref{eq:composite_intro_o2} must contain the same information. However, the fact that $\eqref{eq:orynx}$ is a finite rank vector bundle over $\mathscr{M}_Y \times \mathscr{M}_X$ while \eqref{eq:wuhe20} is a complex of (non-finitely generated) quasi-coherent sheaves explains why $\cat{L}$ is better suited to this kind of problem.

%We leave the details to the aforementioned sequel \cite{??}, but here is a simple example which already shows the nontriviality of viewing correlators as functionals of the moduli of the labels on defects. Consider the following bordism which is a morphism from the empty manifold $\emptyset$ to $O$, where $X$ denotes a matrix factorisation of $V(y) - W(x)$. The value of this diagram is an element
%\[
%\mathcal{Z}(M) \in \mathcal{Z}(O) \cong Jac(V)
%\]
%By virtue of the above discussion, we may think of $\mathcal{Z}(M)$ as a \emph{section} of the trivial vector bundle with fiber $Jac(V)$ over the scheme $\mathscr{M}(W,V)$ which parametrises $1$-morphisms from $W$ to $V$ in $\LG$. If this section is not identically zero, then $V$ is called a \emph{generalised orbifold} of $W$ \cite{??} and it is of great interest to know when this is the case. From what we have said, it is clear that this is essentially a question of the geometry of $\mathscr{M}(W,V)$.

\subsubsection{Semantics of logic}

In forthcoming work we use the results of this paper to define a novel semantics of linear logic, in which formulas are interpreted by tuples $(\md{X}, \nC[x], W)$ where $\md{X}$ is a scheme and $W \in \nC[x]$ is a potential. Proofs in the logic are interpreted by certain correspondences between these pairs. To explain briefly why we are forced to use $\cat{L}$ rather than $\LG$, it is enough to describe the denotation of the Church numeral $\underline{2}$ (see \cite{murfet_ll}).

Let $x$ be a variable of the logic and suppose it has denotation $\llbracket x \rrbracket = (\Spec(\nC), \nC[x], W)$ in the semantics. Let $\mathscr{M} = \mathscr{M}(W,W)$ be the affine scheme parametrising $1$-morphisms $X: W \lto W$ in $\cat{L}$, as discussed above. Then
\begin{gather*}
\llbracket x \multimap x \rrbracket = \big(\Spec(\nC), \nC[x,x'], W(x) - W(x') \big)\,,\\
\llbracket {!}( x \multimap x ) \rrbracket = \big( \mathscr{M}, \nC, 0 \big)\,.
\end{gather*}
The proof $\underline{2}$ of the sequent ${!}(x \multimap x) \vdash x \multimap x$ is roughly speaking a bundle $\llbracket \underline{2} \rrbracket$ over the scheme $\mathscr{M}$ whose fiber over a loop $X$ is the square $X \l X$. More precisely, the denotation is the vector bundle over $\mathscr{M} \times \Spec(\nC[x,x'])$ whose fiber over a point $(X, a)$ is
\be\label{eq:bundledifferentialcliffordproof}
\llbracket \underline{2} \rrbracket_{X,a} = \left( X \l X \right) \otimes_{\nC[x,x']} \nC(a)\,.
\ee
For these definitions to work, it is necessary to be able to compute the coefficients of $X \l X$ as a function of the coefficients of $X$, so we must use $\cat{L}$ in place of $\LG$.

@@

\emph{Acknowledgements.} Thanks to Nils Carqueville for helpful comments on the draft, and to Ingo Runkel and Rafal Suszek for kindly providing Figure \ref{fig:wsh-defect}. %Although the sequel on linear logic is the proper place to fully acknowledge the debt, we cannot neglect to mention the important role that the beautiful ideas in Girard's geometry of interaction program \cite{girard_towards} had in shaping the present work.

@@

Let $R = k[x_1,\ldots,x_n]$ be a polynomial ring and $X$ a finite rank matrix factorisation of $W$ over $R$. Let us fix a homogeneous basis for $X$ so that we may identify $d_X$ with a matrix, and differentiate this matrix entry-wise to obtain the matrix $\lambda_i = \partial_{x_i}(d_X)$. It is easy to check that

\begin{lemma}\label{lemma:homotopy_indept} The odd $R$-linear operator $\lambda_i$ on $X$ is independent, up to homotopy, of the choice of basis.
\end{lemma}

@@

\begin{lemma}\label{lemma:naturalityoflambda} Let $\phi: X \lto X'$ be a morphism of matrix factorisations and write $\lambda_i = \partial_{x_i}(d_X)$ and $\lambda'_i = \partial_{x_i}(d_{X'})$. There is a homotopy $\phi \lambda_i \simeq \lambda'_i \phi$.
\end{lemma}
\begin{proof}
From $d_{X'} \phi = \phi d_X$ we deduce
\[
\partial_{x_i}( d_{X'} ) \phi + d_{X'} \partial_{x_i}( \phi ) = \partial_{x_i}( \phi ) d_X + \phi \partial_{x_i}( d_X )
\]
and hence
\[
\partial_{x_i}( d_{X'} ) \phi - \phi \partial_{x_i}( d_X ) = \partial_{x_i}(\phi) d_X - d_{X'} \partial_{x_i}(\phi)
\]
as claimed.
\end{proof}

@@

The cut operation in $\cat{L}$ produces objects which are representations of Clifford algebras. We briefly recall the basic theory of these algebras and their modules; for a full discussion see \cite{friedrich}. For $n \ge 0$ the Clifford algebra $C_n$ is the associative $\mathbb{Z}_2$-graded $k$-algebra generated by odd elements $\gamma_1,\ldots,\gamma_n$ and $\gamma_1^\dagger, \ldots, \gamma_n^\dagger$ subject to the following relations
\be\label{eq:clifford_relations}
[\gamma_i, \gamma_j] = 0, \qquad [\gamma_i^\dagger, \gamma_j^\dagger] = 0, \qquad [\gamma_i, \gamma_j^\dagger] = \delta_{ij}\,.
\ee
for all $1 \le i, j \le n$. In this paper all our commutators are graded, $[a,b] = ab - (-1)^{|a||b|} ba$. We set $C_0 = k$. These Clifford algebras have essentially only one nontrivial representation, sometimes called the \emph{spinor representation} \cite[p.14]{friedrich}. To define this representation we use the $\nZ_2$-graded $k$-module
\[
F_n = k \theta_1 \oplus \cdots \oplus k \theta_n
\]
where the $\theta_i$ are formal variables of odd degree, and set
\be
S_n = \bigwedge F_n = \bigwedge\big( k \theta_1 \oplus \cdots \oplus k \theta_n \big)\,.
\ee

@@

In particular this means that every $\mathbb{Z}_2$-graded $C_n$-module is isomorphic to $S_n \otimes_k V$ for some $\mathbb{Z}_2$-graded $k$-module $V$. Using the isomorphism of $k$-modules
\[
F_m \oplus F_n \cong F_{m+n}
\]
which maps the ordered basis $\theta_1,\ldots,\theta_m$ of $F_m$ to the first $m$ basis elements of $F_{m+n}$ in their given order, and the ordered basis of $F_n$ to the last $n$ basis elements, we define
\[
S_m \otimes_k S_n = \bigwedge F_m \otimes_k \bigwedge F_n \cong \bigwedge F_{m+n} = S_{m+n}\,.
\]
From this we deduce an isomorphism of algebras
\begin{equation}\label{eq:algebra_A_additive}
C_m \otimes_k C_n \cong C_{m+n}\,.
\end{equation}
Given $m, n \ge 0$ we introduce notation for the $C_m$-$C_n$-bimodule
\begin{equation}
S_{m,n} = S_m \otimes_k S_n^* \cong \Hom_k(S_n,S_m)
\end{equation}
where $S_n^* = \Hom_k(S_n, k)$. There is an isomorphism of $C_l$-$C_n$-bimodules
\begin{equation}\label{eq:isosbimodule}
S_{l,m} \otimes_{C_m} S_{m,n} = S_l \otimes_k S_m^* \otimes_{C_m} S_m \otimes_k S_n^* \cong S_{l,n}\,.
\end{equation}

@@

\section{The Clifford thickening}\label{section:cliff_thick}

@@

More concretely $\cat{T}^\bullet$ is the category with objects all pairs $(X,n)$ of an integer $n \ge 0$ and an object $X \in \cat{T}$ together with a left $C_n$-action. A morphism $\alpha: (X,n) \lto (Y,m)$ is a morphism of $C_m$-modules $\alpha: S_{m,n} \otimes_{C_n} X \lto Y$. Composition of a pair of morphisms
\begin{equation}\label{eq:comp_chain0}
\xymatrix{
(X,n) \ar[r]^-{\alpha} & (Y,m) \ar[r]^-{\beta} & (Z,l)
}
\end{equation}
is given by the morphism of $C_l$-modules
\begin{equation}\label{eq:comp_chain1}
\xymatrix@C+1pc{
S_{l, n} \otimes_{C_n} X \ar[r]^-{\cong} & (S_{l,m} \otimes_{C_m} S_{m,n}) \otimes_{C_n} X \ar[d]^-{\cong}\\
& S_{l,m} \otimes_{C_m} ( S_{m,n} \otimes_{C_n} X ) \ar[r]^-{ 1 \otimes \alpha } & S_{l,m} \otimes_{C_m} Y \ar[r]^-{\beta} & Z\,.
}
\end{equation}
There is a canonical functor $p: \cat{T}^\bullet \lto \mathbb{N}$ defined by $p(X,n) = n$, which defines a cofibered category over $\mathbb{N}$ by the definition of the Grothendieck construction. This definition has the benefit of being derived from a general construction, but there is an alternative description of morphisms which is less cumbersome:

@@

In this section we define a superbicategory without units $\cat{L}$. Later we will prove that this bicategory is equivalent to $\LG$, although this will not be apparent at first. The objects of $\cat{L}$ are the same as $\LG$, namely potentials $(x,W)$.

\begin{definition} Given potentials $(x,W)$ and $(y,V)$ we define
\be\label{eq:defnluv}
\cat{L}(W,V) = \Big( \hmf( k[x,y], V - W )^{\omega} \Big)^{\bullet}\,,
\ee
where $(-)^\omega$ denotes the idempotent completion (see Remark \ref{remark:idempotent_completion}) and $(-)^{\bullet}$ is the Clifford thickening (see Section \ref{section:cliff_thick}).
\end{definition}

Thus a $1$-morphism $W \lto V$ in $\cat{L}$ is a finite rank matrix factorisation $X$ of $V(y) - W(x)$ together with an idempotent endomorphism $e$ and a family of odd operators $\gamma_i, \gamma_i^\dagger$ for $1 \le i \le a$ satisfying Clifford relations and the equations (all up to homotopy)
\begin{gather*}
\gamma_i e = \gamma_i = e \gamma_i\,, \\
\gamma_i^\dagger e = \gamma_i^\dagger = e \gamma_i^\dagger\,.
\end{gather*}
Most of the time we only care about $1$-morphisms where $e = 1$, and let us only describe $2$-morphisms in this case. For matrix factorisations $X,Y$ of $V(y) - W(x)$ with respective Clifford actions $\{ \gamma_i, \gamma_i^\dagger \}_{i=1}^a$ and $\{ \rho_j, \rho_j^\dagger \}_{j=1}^b$ the $2$-morphisms $\phi: (X,a) \lto (Y,b)$ in $\cat{L}$ are by Lemma \ref{lemma:equivcliffordthick} in bijection with morphisms of matrix factorisations $\phi: X \lto Y$ satisfying
\[
\phi \gamma_i = 0\,, \qquad \rho_j^\dagger \phi = 0\,, \qquad \forall 1 \le i \le a, 1 \le j \le b\,.
\]
Observe that the homotopy category $\LG(W,V)_{fin}$ of finite rank matrix factorisations sits fully faithfully inside $\cat{L}(W,V)$ as the subcategory where all idempotents are identities and there are no Clifford actions, and by Lemma \ref{lemma:embedcliffordthick} the inclusion of this subcategory into $\cat{L}(W,V)$ is an idempotent completion.

 The first aim of this section is to define, for any object $(z,U)$ of $\LG$, a functor
\begin{gather*}
\cat{L}(V,U) \otimes_k \cat{L}(W,V) \lto \cat{L}(W, U)\,\\
(Y,X) \mapsto Y \l X
\end{gather*}
which we call the \textsl{cut functor}. The cut operation is defined on matrix factorisations $X$ of $V - W$ and $Y$ of $U - V$ as follows, assuming $y = (y_1,\ldots,y_m)$ and writing
\be\label{eq:defnjacobian}
J_V = k[y] / ( \partial_{y_1} V, \ldots, \partial_{y_m} V )\,.
\ee

@@

\begin{remark}\label{remark:inflation} If we choose a $k$-basis for $J_V$ then multiplication by $y_i$, as a $k$-linear operator $J_V \lto J_V$, gives a matrix $[y_i]$ over $k$.
